---
title: "01. EXPLORATORY ANALYSIS"
author: "Gary Nguyen"
params:
  data_folder: /Users/nguyenh/Desktop/personal_projects/Health_Insurance_Marketplace/data
output:
  output: rmarkdown::github_document
    df_print: paged
---

## 1. INITIAL SETTINGS

Data descriptions:
- Benefit Name: the name of the benefit (837 unique benefits)
- Business Year: year (2014, 2015 and 2016)
- CoinsInnTier1, CoinsInnTier2: coinsurance % for Tier 1 and 2, respectively
- CoinsOutofNet
- CopayInnTier1, CopayInnTier2
- CopayOutofNet
- EHBVarReason
- Exclusions
- Explanation
- ImportDate
- IsCovered
- IsEHB
- IsExclFromInnMOOP
- IsExclFromOonMOOP
- IsStateMandate
- IsSubjToDedTier1, IsSubjToDedTier2
- IssuerId, IssuerId2
- LimitQty
- LimitUnit
- MinimumStay
- PlanId
- QuantLimitOnSvc
- RowNumber
- SourceName
- StandardComponentId
- StateCode, StateCode2
- VersionNum

```{r loading_libraries, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
if(!require(ggplot2, quietly = TRUE)) install.packages('ggplot2')
if(!require(readr, quietly = TRUE)) install.packages('readr')
if(!require(dplyr, quietly = TRUE)) install.packages('dplyr')
if(!require(choroplethr, quietly = TRUE)) install.packages('choroplethr')
if(!require(extrafont, quietly = TRUE)) install.packages('extrafont')
if(!require(extrafontdb, quietly = TRUE)) install.packages('extrafontdb')
if(!require(RColorBrewer, quietly = TRUE)) install.packages('RColorBrewer')
if(!require(maps, quietly = TRUE)) install.packages('maps')
if(!require(scales, quietly = TRUE)) install.packages('scales')
if(!require(mapproj, quietly = TRUE)) install.packages('mapproj')
if(!require(ggthemes, quietly = TRUE)) install.packages('ggthemes')
```

```{r loading_packages, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
pkg <- c("ggplot2",
         "readr",
         "dplyr",
         "choroplethr",
         "extrafont",
         "extrafontdb",
         "RColorBrewer",
         "knitr",
         "tidyr",
         "stringr",
         "lubridate",
         "purrr",
         "DT",
         "tidytext",
         "textstem",
         "tm",
         "splitstackshape",
         "text2vec",
         "reshape",
         "readr",
         "zoo",
         "keras",
         "ROCR",
         "caret",
         "maps",
         "ggthemes",
         "scales",
         "mapproj")
invisible(lapply(pkg, library, character.only = TRUE))
options(warn=0)
```

```{r helper_data_states}
states_abbr = c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA",
                "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME",
                "MI", "MN", "MO", "MS",  "MT", "NC", "ND", "NE", "NH", "NJ", "NM",
                "NV", "NY", "OH", "OK", "OR", "PA", "PR", "RI", "SC", "SD", "TN",
                "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY")
                      
states_full = c("alaska","alabama","arkansas","arizona","california","colorado",
                "connecticut","district of columbia","delaware","florida","georgia",
                "hawaii","iowa","idaho","illinois","indiana","kansas","kentucky",
                "louisiana","massachusetts","maryland","maine","michigan","minnesota",
                "missouri","mississippi","montana","north carolina","north dakota",
                "nebraska","new hampshire","new jersey","new mexico","nevada",
                "new york","ohio","oklahoma","oregon","pennsylvania","puerto rico",
                "rhode island","south carolina","south dakota","tennessee","texas",
                "utah","virginia","vermont","washington","wisconsin",
                "west virginia","wyoming")

names(states_full) <- states_abbr
```

```{r loading_data}
data_file_path <- file.path(params$data_folder, "BenefitsCostSharing.csv")
benefits_cost_sharing <- readr::read_csv(data_file_path)
```

## 2. LOOKING AT SOME BENEFITS {.tabset}

### 2A. INFUSION THERAPY

```{r}
infusion_therapy <- benefits_cost_sharing %>% 
  filter(BenefitName == "Infusion Therapy")

infusion_therapy <- infusion_therapy %>% 
  mutate(CoinsInnTier1_processed = as.integer(stringr::str_extract(CoinsInnTier1, "\\d+")),
         CoinsInnTier1_processed = replace(CoinsInnTier1_processed, is.na(CoinsInnTier1_processed), 0),
         CoinsInnTier1_processed = CoinsInnTier1_processed/100)
```

```{r q_plot}
qplot(CoinsInnTier1_processed, 
      data = infusion_therapy, 
      geom = "bar", 
      xlab = "Co-pay", 
      ylab = "Number of Plans", 
      main = "Health Insurers with Infusion Co-Pays") +
  theme(text = element_text(size = 16, family = "Garamond"))
```

```{r}
infusion_to_map <- infusion_therapy %>% 
  group_by(StateCode) %>% 
  summarize(value = mean(CoinsInnTier1_processed)) %>% 
  mutate(region = if_else(StateCode %in% names(states_full), states_full[StateCode], StateCode))
```

```{r}
choro <- infusion_to_map %>% 
  StateChoropleth$new()
choro$title = "Average CoPay for Infusion Services"
choro$set_num_colors(1)
myPalette <- colorRampPalette(brewer.pal(9, "Reds"))
choro$ggplot_polygon = geom_polygon(aes(fill = value), color = NA)
choro$ggplot_scale = scale_fill_gradientn(name = "CoPay", colours = myPalette(9))
choro$render()
```

Looking at states with terrible co-pays

```{r}
x <- infusion_therapy %>% 
  filter(BenefitName == "Infusion Therapy",
         CoinsInnTier1_processed >= 0.5)
qplot(BusinessYear,
      data = x,
      geom = "bar", 
      fill = StateCode, 
      xlab = "Business Year", 
      ylab="Number of Plans", 
      main = "Health Insurers with Infusion Co-Pays of 50% or more") + 
theme(text = element_text(size = 8, family = "Garamond"))
```


### 2B. MENTAL HEALTH OFFICE VISITS ("Abortion for Which Public Funding is Prohibited")

```{r}
abortion <- benefits_cost_sharing %>% 
  filter(BenefitName == "Abortion for Which Public Funding is Prohibited") %>% 
  mutate(CoinsInnTier1_processed = as.integer(stringr::str_extract(CoinsInnTier1, "\\d+")),
         CoinsInnTier1_processed = replace(CoinsInnTier1_processed, is.na(CoinsInnTier1_processed), 0),
         CoinsInnTier1_processed = CoinsInnTier1_processed/100)

abortion_to_map <- abortion %>% 
  group_by(StateCode) %>% 
  summarize(value = mean(CoinsInnTier1_processed)) %>% 
  mutate(region = if_else(StateCode %in% names(states_full), states_full[StateCode], StateCode))
```

```{r}
choro <- abortion_to_map %>% 
  StateChoropleth$new()
choro$title = "Average CoPay for Abortions"
choro$set_num_colors(1)
myPalette <- colorRampPalette(brewer.pal(9, "Reds"))
choro$ggplot_polygon = geom_polygon(aes(fill = value), color = NA)
choro$ggplot_scale = scale_fill_gradientn(name = "CoPay", colours = myPalette(9))
choro$render()
```

```{r}
x <- abortion %>% 
  filter(BenefitName == "Abortion for Which Public Funding is Prohibited",
         CoinsInnTier1_processed >= 0.1)
qplot(BusinessYear,
      data = x,
      geom = "bar", 
      fill = StateCode, 
      xlab = "Business Year", 
      ylab="Number of Plans", 
      main = "Health Insurers with Abortion Co-Pays of 50% or more") + 
theme(text = element_text(size = 8, family = "Garamond"))
```

## 3. LOOKING AT RATES {.tabset}

```{r}
rate_file_path <- file.path(params$data_folder, "Rate.csv")
rate <- readr::read_csv(rate_file_path)

plan_attribute_file_path <- file.path(params$data_folder, "PlanAttributes.csv")
plan_attribute <- readr::read_csv(plan_attribute_file_path)
```

```{r}
rate_individual_2015 <- rate %>% 
  filter(BusinessYear == "2015",
         Age != "Family Option",
         IndividualRate < "9000") %>% 
  select(c(BusinessYear:IndividualTobaccoRate))
```

```{r}
#first step is to pull 2015 Gold Silver Bronze ACA plans. 
plan_attribute_processed <- plan_attribute %>% 
  select(c(115, 104, 162, 166, 170)) %>% 
  mutate(PlanId = substr(PlanId, 1, 14),
         TEHBInnTier1IndividualMOOP = str_replace_all(TEHBInnTier1IndividualMOOP, ",", ""),
         TEHBInnTier1IndividualMOOP = as.numeric(str_replace_all(TEHBInnTier1IndividualMOOP, "\\$", "")),
         TEHBInnTier2IndividualMOOP = str_replace_all(TEHBInnTier2IndividualMOOP, ",", ""),
         TEHBInnTier2IndividualMOOP = as.numeric(str_replace_all(TEHBInnTier2IndividualMOOP, "\\$", "")),
         TEHBOutOfNetIndividualMOOP = str_replace_all(TEHBOutOfNetIndividualMOOP, ",", ""),
         TEHBOutOfNetIndividualMOOP = as.numeric(str_replace_all(TEHBOutOfNetIndividualMOOP, "\\$", "")))

#Finalize the benefit table to use in the matching process
benefit_grouped <- plan_attribute_processed %>% 
  group_by(PlanId, MetalLevel) %>%
  summarise(innettier1moop = mean(TEHBInnTier1IndividualMOOP, na.rm = TRUE),
            innettier2moop = mean(TEHBInnTier2IndividualMOOP, na.rm = TRUE),
            outnetmoop = mean(TEHBOutOfNetIndividualMOOP, na.rm = TRUE))
```


### 3A. NUMBER OF PLANS BY STATE

```{r}
to_map <- rate_individual_2015 %>% 
  group_by(StateCode) %>%
  summarize(num_carriers = length(unique(IssuerId)), 
            num_plan_offered = length(unique(PlanId)),
            mean_ind_rate = mean(IndividualRate),
            median_ind_rate = median(IndividualRate)) %>%
  arrange(desc(num_plan_offered))
```

```{r}
to_map <- to_map %>% 
  mutate(num_range_carriers = if_else(num_carriers < 15, "(0,15)", 
                                      if_else(num_carriers >= 25, "[25,35)",
                                              "[15,25)")))
         
a <- ggplot(to_map, aes(x = reorder(StateCode, num_plan_offered), y = num_plan_offered)) +
  geom_bar(aes(fill = num_range_carriers), stat = "identity") + 
  coord_flip() +
  ggtitle("Carriers vs. Plans Available By State")+
  labs(x = "State", y = "Plans Available")
print(a)
```

### 3B. MEDIAN MONTHLY PREMIUM

```{r}
to_map$region <- tolower(state.name[match(to_map$StateCode, state.abb)])

#load US map
us_state_map = map_data('state')
statename <- us_state_map %>% 
  group_by(region) %>% 
  summarise(long = mean(long), 
            lat = mean(lat))

mapdata <- left_join(to_map, 
                     us_state_map, 
                     by="region")

p <- ggplot() + 
  geom_polygon(data = mapdata, 
               aes(x = long, y = lat, group = group, fill = mapdata$median_ind_rate), colour = "white") +
  scale_fill_continuous(low = "thistle2", high = "darkred", guide = "colorbar") + 
  theme_bw()+
  labs(fill = "Median Premium $/mon", title = "Median Monthly Premium Distribution", x = "", y = "") +
  scale_y_continuous(breaks = c()) + 
  scale_x_continuous(breaks = c()) + 
  theme(panel.border = element_blank()) +
  geom_text(data = statename, aes(x = long, y = lat, label = region),  na.rm = T, size = 2) +
  coord_map()

print(p)
```

### 3C. NUMBER OF PLANS BY STATE








